Centrality and Activity for Envolving Networks

### 2013-06-24 network-science/paper.tex

\noindent \textbf{Keywords:} social network, dynamic graph, vertex activity,
vertex centrality, online algorithm.
\end{abstract}
% \tableofcontents
\newpage

@@

\subsection{Desiderata}
\definition $\mathit{Act}(G_t,v)$ returns the activity of node $v$ in the evolving graph in
time $t$. Then $\mathit{Act}(G_t,v)$ should be a deterministic function of $v$ and
$E_t$.
\begin{itemize}
\item \textbf{(normalization)}:

@@

\item \textbf{(initialization)}:

@@

\begin{equation}
\begin{split}
\int_{t_i}^{t_i+\eta w(t_i)} a(G_t,v) dt &= \int_{t_i}^{t_i+\eta w(t_i)} e^{\frac{1}{t-t_i-\eta w(t_i)}} dt \\
&= \int_{-\eta w(t_i)}^{0} e^{\frac{1}{t}} dt.
\end{split}
\end{equation}

Notice that $e^{1/t}<1(t<0)$, so it's easy for this direction
\begin{equation}
\int_{-\eta w(t_i)}^{0} e^{\frac{1}{t}} dt< \eta w(t_i).
\end{equation}

Define $f(x)=\int_{-x}^{0} e^{1/t} dt - \frac{x}{10}$, from $f'(x)=e^{-1/x}-\frac{1}{10}$, at least we can see $f(x)$ increases when $x\ge 1$. So $f(x)\ge f(1)\approx 0.05>0$.

Now we have the other direction: 
\begin{equation}
\int_{-\eta w(t_i)}^{0} e^{\frac{1}{t}} dt>\frac{\eta w(t_i)}{10},
\end{equation}
as long as $\eta w(t_i)\ge 1$.

@@

One can verify that
\begin{equation}
\int_{t_i}^{t_j} a(G_t,v) dt< \eta \sum_{k=i}^j w(t_k),
\end{equation}
easily since we still have $\mu^{1/z(t)}<1$.

As for the other direction, notice that at time $t_{k+1}$, there are two cases. One is that $z(t_{k+1})$ has increased equal or above $z(t_k)$, namely we can find some $t_k^*\in (t_k, t_{k+1})$, where $t_k^*$ is defined to satisfy that $z(t_k)=z(t_k^*)$ for all $k$. So we have
\begin{equation}
\int_{t_k}^{t_{k+1}} \mu^{\frac{1}{z(t)}} dt \ge \int_{t_k}^{t_k^*} \mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}dt.
\end{equation}
The other case is that $z(t_{k+1})$ is still less than $z(t_{k})$, in other words, we will find $t_k^*>t_{k+1}$.

@@

we have:
\begin{equation}
\begin{split}
\int_{t_k}^{t_{k+2}} \mu^{\frac{1}{z(t)}} dt &\ge \int_{t_k}^{t_{k+1}^*} \mu^{\frac{1}{z(t)}} dt \\
&= \int_{t_k}^{t_{k+1}} \mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}  dt + \int_{t_{k+1}}^{t_{k+1}^*} \mu^{\frac{1}{z(t_{k+1})+t-t_{k+1}}} -  \mu^{\frac{1}{z(t_{k+1})}} dt + (t_{k+1}^*-t_{k+1})\mu^{\frac{1}{z(t_{k+1})}}\\
&\ge \int_{t_k}^{t_{k}^*} \mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}dt + \int_{t_{k+1}}^{t_{k+1}^*} \mu^{\frac{1}{z(t_{k+1})+t-t_{k+1}}} - \mu^{\frac{1}{z(t_{k+1})}}dt,
\end{split}
\end{equation}
by an easy induction, one can split a "phase" into integrals in the form 
\begin{equation}
\int_{t_k}^{t_{k}^*} \mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}dt
\end{equation}
 Since $ \mu^{1/z}$ decreases as $z$ becomes larger, we can reduce it to the base case now.

@@

\begin{algorithm}[htbp]
\caption{\textsc{Add Edge}($v_i,v_j,t,w$)}
\label{alg-add}
\textbf{Time Complexity:} $O(\log |V|)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\Require $1\leq i,j\leq |V|$ and $t\geq \mathit{currentTime}$ and $w>0$.
\Ensure $G_t$ is up-to-date.
\State $\mathit{currentTime}\gets t$;
\State $z_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.z$;
\State $t_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.t$;
\State $s_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.s$;
\State $z_1\gets z_1+\mathit{currentTime}-t_1$;
\If {$z_1>0$} \State $z_1\gets 0$;\EndIf
\State $z_1\gets z_1-\eta\cdot w$;
\State $z_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.z$;
\State $t_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.t$;
\State $z_2\gets z_2+\mathit{currentTime}-t_2$;
\State $s_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.s$;
\If {$z_2>0$} \State $z_2\gets 0$;\EndIf
\State $z_2\gets z_2-\eta\cdot w$;
\State append ($j$, $t$, $w$, $s_2+w$, $z_2$) to \textit{edgeList}[$i$];
\State append ($i$, $t$, $w$, $s_1+w$, $z_1$) to \textit{edgeList}[$j$];
\If{ binary search on \textit{edgeListReduced}[$i$] did not find $j$}
\State insert $j$ to \textit{edgeListReduced}[$i$] in appropriate position;
\State binary search $i$ on \textit{edgeListReduced}[$j$];
\State insert $i$ to \textit{edgeListReduced}[$j$] in appropriate position;
\EndIf
\State $p\gets \Call{FindRoot}{i,\mathit{currentTime}}$;
\State $q\gets \Call{FindRoot}{j,\mathit{currentTime}}$;
\If{$p\neq q$}
\State \Call{Merge}{$p$,$q$}
\EndIf
\end{algorithmic}
\end{algorithm}

@@

\section{Experiment}
\subsection{Data Generation} 
Our data generation is base on real social networks provided by \cite{database1,
database2}. Unfortunately, those databases do not contain evolving history
recorded data, so we simulate the evolving history by use the
authentic graph as the final state of the tested dynamic graph.
We give a random permutation to the edges, in a way such that there will always
be just one component -- known as the giant component. Because the final state
is from real world data, it is ensured 
the evolving graph generated by our model will satisfy all the property that we
discussed in section 2.2. In addition, Kumar introduced another way to generate
artificial data of evolving social networks\cite{kumar2010structure}.

@@

The centrality trace graph for an representative node, when $s$:

>>>>

\noindent \textbf{Keywords:} social network, dynamic graph, vertex activity,
centrality, online algorithm.
\end{abstract}
% \tableofcontents
\newpage

@@

\subsection{Desiderata}
\definition $\mathit{Act}(G_t,v)$ returns the activity of node $v$ in the evolving graph in
time $t$. Then $\mathit{Act}(G_t,v)$ should be a deterministic function of $v$ and
$E_t$.
\begin{itemize}
\item \textbf{normalization}:

@@

\item \textbf{initialization}:

@@

\begin{equation}
\begin{split}
\int_{t_i}^{t_i+\eta w(t_i)} a(G_t,v) \mathrm{d}t &= \int_{t_i}^{t_i+\eta
w(t_i)} e^{\frac{1}{t-t_i-\eta w(t_i)}} \mathrm{d}t \\
&= \int_{-\eta w(t_i)}^{0} e^{\frac{1}{t}} \mathrm{d}t.
\end{split}
\end{equation}

Notice that $e^{1/t}<1(t<0)$, so it's easy for this direction
\begin{equation}
\int_{-\eta w(t_i)}^{0} e^{\frac{1}{t}} \mathrm{d}t< \eta w(t_i).
\end{equation}

Define $f(x)=\int_{-x}^{0} e^{1/t} \mathrm{d}t - \frac{x}{10}$, from $f'(x)=e^{-1/x}-\frac{1}{10}$, at least we can see $f(x)$ increases when $x\ge 1$. So $f(x)\ge f(1)\approx 0.05>0$.

Now we have the other direction: 
\begin{equation}
\int_{-\eta w(t_i)}^{0} e^{\frac{1}{t}} \mathrm{d}t>\frac{\eta w(t_i)}{10},
\end{equation}
as long as $\eta w(t_i)\ge 1$.

@@

One can verify that
\begin{equation}
\int_{t_i}^{t_j} a(G_t,v) \mathrm{d}t< \eta \sum_{k=i}^j w(t_k),
\end{equation}
easily since we still have $\mu^{1/z(t)}<1$.

As for the other direction, notice that at time $t_{k+1}$, there are two cases. One is that $z(t_{k+1})$ has increased equal or above $z(t_k)$, namely we can find some $t_k^*\in (t_k, t_{k+1})$, where $t_k^*$ is defined to satisfy that $z(t_k)=z(t_k^*)$ for all $k$. So we have
\begin{equation}
\int_{t_k}^{t_{k+1}} \mu^{\frac{1}{z(t)}} \mathrm{d}t \ge \int_{t_k}^{t_k^*}
\mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}\mathrm{d}t.
\end{equation}
The other case is that $z(t_{k+1})$ is still less than $z(t_{k})$, in other words, we will find $t_k^*>t_{k+1}$.

@@

we have:
\begin{equation}
\begin{split}
&\int_{t_k}^{t_{k+2}} \mu^{\frac{1}{z(t)}} \mathrm{d}t\\
\ge{}& \int_{t_k}^{t_{k+1}^*} \mu^{\frac{1}{z(t)}} \mathrm{d}t \\
={}&\int_{t_k}^{t_{k+1}} \mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}
\mathrm{d}t+ \int_{t_{k+1}}^{t_{k+1}^*} \mu^{\frac{1}{z(t_{k+1})+t-t_{k+1}}}\\& -
\mu^{\frac{1}{z(t_{k+1})}} \mathrm{d}t + (t_{k+1}^*-t_{k+1})\mu^{\frac{1}{z(t_{k+1})}}\\
\ge{}& \int_{t_k}^{t_{k}^*} \mu^{\frac{1}{z(t_k)+t-t_k}} -
\mu^{\frac{1}{z(t_{k})}}\mathrm{d}t + \int_{t_{k+1}}^{t_{k+1}^*}
\mu^{\frac{1}{z(t_{k+1})+t-t_{k+1}}} - \mu^{\frac{1}{z(t_{k+1})}}\mathrm{d}t,
\end{split}
\end{equation}
by an easy induction, one can split a "phase" into integrals in the form 
\begin{equation}
\int_{t_k}^{t_{k}^*} \mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}\mathrm{d}t
\end{equation}
 Since $ \mu^{1/z}$ decreases as $z$ becomes larger, we can reduce it to the base case now.

@@

\begin{algorithm}[htbp]
\caption{\textsc{Add Edge}($v_i,v_j,t,w$)}
\label{alg-add}
\textbf{Time Complexity:} $O(\log |V|)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\Require $1\leq i,j\leq |V|$ and $t\geq \mathit{currentTime}$ and $w>0$.
\Ensure $G_t$ is up-to-date.
\State $\mathit{currentTime}\gets t$;
\State $z_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.z$;
\State $t_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.t$;
\State $s_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.s$;
\State $z_1\gets z_1+\mathit{currentTime}-t_1$;
\If {$z_1>0$} \State $z_1\gets 0$;\EndIf
\State $z_1\gets z_1-\eta\cdot w$;
\State $z_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.z$;
\State $t_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.t$;
\State $z_2\gets z_2+\mathit{currentTime}-t_2$;
\State $s_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.s$;
\If {$z_2>0$} \State $z_2\gets 0$;\EndIf
\State $z_2\gets z_2-\eta\cdot w$;
\State append ($j$, $t$, $w$, $s_2+w$, $z_2$) to \textit{edgeList}[$i$];
\State append ($i$, $t$, $w$, $s_1+w$, $z_1$) to \textit{edgeList}[$j$];
\If{ binary search on \textit{edgeListReduced}[$i$] did not find $j$}
\State insert $j$ to \textit{edgeListReduced}[$i$] in appropriate position;
\State binary search $i$ on \textit{edgeListReduced}[$j$];
\State insert $i$ to \textit{edgeListReduced}[$j$] in appropriate position;
\EndIf
\State $p\gets \Call{Find Root}{i,\mathit{currentTime}}$;
\State $q\gets \Call{Find Root}{j,\mathit{currentTime}}$;
\If{$p\neq q$}
\State \Call{Merge}{$p$,$q$}
\EndIf
\end{algorithmic}
\end{algorithm}

@@

\section{Experiment}
\subsection{Data Generation} 
Our data generation is base on real social networks provided by \cite{database1,
database2}. Unfortunately, those databases do not contain evolving history
recorded data, so we simulate the evolving history by use the
authentic graph as the final state of the tested dynamic graph.
We give a random permutation to the edges, in a way such that:
\begin{itemize}
\item There will always
be just one component -- known as the giant component. We never add an edge out
of the giant component, so the evolving of the network seems to have increasing
vertices joining to the network.
\item The time delay of two adjacent new edges is inversely proportional to the
size of the giant component. So the average activity of any indivitual node do
not change tremendously.
\end{itemize}
Because the final state
is from real world data, it is ensured 
the evolving graph generated by our model will satisfy all the property that we
discussed in section 2.2. In addition, Kumar introduced how to generate
an artificial evolving social networks\cite{kumar2010structure}.

@@

The centrality trace graph for some representative node, when $\mu=4096$ and
$\eta=1000$:
\begin{center} \begin{tikzpicture}[y=.4cm, x=.5cm, font=\sffamily]
	\draw (0,0) -- coordinate (x axis mid) (21, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 11);
	\node[below=0.5cm] at (x axis mid) {\scriptsize Time ($\times10^3$)};
	\node[rotate=90, left=0.8cm] at (0, 10) {\scriptsize Magnitude of
	Centrality};
	\foreach \x in {1,2,...,20}
	\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x};
		\foreach \y in {1,2,...,11}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	\draw[color=red!70!black,line width=1pt] plot file {exp-g3-1};
	\draw[color=blue!70!black] plot file {exp-g3-2};
	\begin{scope}[shift={(2,8)}]
	\draw[color=red!70!black,line width=1pt] (0,0) -- (2,0)
		node[right]{\scriptsize evolving centrality};
	\draw[color=blue!50!black] (0,1) -- (2,1)
		node[right]{\scriptsize static centrality};
	\end{scope}
\end{tikzpicture} \end{center}

\begin{center} \begin{tikzpicture}[y=.4cm, x=.5cm, font=\sffamily]
	\draw (0,0) -- coordinate (x axis mid) (21, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 11);
	\node[below=0.5cm] at (x axis mid) {\scriptsize Time ($\times10^3$)};
	\node[rotate=90, left=0.8cm] at (0, 10) {\scriptsize Magnitude of
	Centrality};
	\foreach \x in {1,2,...,20}
	\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x};
		\foreach \y in {1,2,...,11}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	\draw[color=red!70!black,line width=1pt] plot file {exp-g9-1};
	\draw[color=blue!70!black] plot file {exp-g9-2};
	\begin{scope}[shift={(2,8)}]
	\draw[color=red!70!black,line width=1pt] (0,0) -- (2,0)
		node[right]{\scriptsize evolving centrality};
	\draw[color=blue!50!black] (0,1) -- (2,1)
		node[right]{\scriptsize static centrality};
	\end{scope}
\end{tikzpicture} \end{center}

The centrality trace graph for some representative node, when $\mu=2^{20}$ and
$\eta=2000$:
\begin{center} \begin{tikzpicture}[y=.4cm, x=.5cm, font=\sffamily]
	\draw (0,0) -- coordinate (x axis mid) (21, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 11);
	\node[below=0.5cm] at (x axis mid) {\scriptsize Time ($\times10^3$)};
	\node[rotate=90, left=0.8cm] at (0, 10) {\scriptsize Magnitude of
	Centrality};
	\foreach \x in {1,2,...,20}
	\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x};
		\foreach \y in {1,2,...,11}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	\draw[color=red!70!black,line width=1pt] plot file {exp-g8-1};
	\draw[color=blue!70!black] plot file {exp-g8-2};
	\begin{scope}[shift={(2,8)}]
	\draw[color=red!70!black,line width=1pt] (0,0) -- (2,0)
		node[right]{\scriptsize evolving centrality};
	\draw[color=blue!50!black] (0,1) -- (2,1)
		node[right]{\scriptsize static centrality};
	\end{scope}
\end{tikzpicture} \end{center}
\begin{center} \begin{tikzpicture}[y=.4cm, x=.5cm, font=\sffamily]
	\draw (0,0) -- coordinate (x axis mid) (21, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 11);
	\node[below=0.5cm] at (x axis mid) {\scriptsize Time ($\times10^3$)};
	\node[rotate=90, left=0.8cm] at (0, 10) {\scriptsize Magnitude of
	Centrality};
	\foreach \x in {1,2,...,20}
	\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x};
		\foreach \y in {1,2,...,11}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	\draw[color=red!70!black,line width=1pt] plot file {exp-g11-1};
	\draw[color=blue!70!black] plot file {exp-g11-2};
	\begin{scope}[shift={(2,8)}]
	\draw[color=red!70!black,line width=1pt] (0,0) -- (2,0)
		node[right]{\scriptsize evolving centrality};
	\draw[color=blue!50!black] (0,1) -- (2,1)
		node[right]{\scriptsize static centrality};
	\end{scope}
\end{tikzpicture} \end{center}
